---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# norm

<!-- badges: start -->
<!-- badges: end -->

The goal of norm is to normalize the data using Min-Max Normalization scale.

## Installation

You can install the development version of norm like so:

``` r
devtools::install_github("stat545ubc-2022/assignment-b1-and-b2-slinger0225", ref = "0.1.0")
```

Then, load the norm() function

```{r}
library(norm)
```


## Example

Those are basic examples which shows you how to solve a common problem:

## Normalize a column

The first example is demonstrated using *height_range_id* and *diameter_level* column from *vancouver_trees* dataset.

After normalization, as you can see below, the range of both column becomes [0, 1]
```{r}
library(datateachr)
height_norm <- norm(vancouver_trees$height_range_id)
diameter_norm <- norm(vancouver_trees$diameter)
height_norm_range <- range(height_norm, na.rm = TRUE)
cat("Range of height_range_id is:[", height_norm_range[1], ",",  height_norm_range[2], "]")
diameter_norm_range <-range(diameter_norm, na.rm = TRUE)
cat("Range of height_range_id is:[", diameter_norm_range[1], ",",  diameter_norm_range[2], "]")
```

### Error column

To calculate the normalized data, the data itself should be numeric. 

In *vancouver_trees* data, *neighbourhood_name* is a character column.
```{r}
head(vancouver_trees$neighbourhood_name)
```


Apply norm() to it will produce an error:
```{r, error = TRUE}
neighbourhood_norm <- norm(vancouver_trees$neighbourhood_name)
```

## Normalize a vector

### Normal vector

```{r}
data <- c(1,2,3,4)
```

After normalization, each of the data element become:
```{r}
vector_norm <- norm(data)
print(vector_norm)
```

### Constant vector
```{r}
data_constant <- c(4, 4, 4, 4)
```


Constant data can't be normalized. According to our function's definition, we should return the original data:
```{r}
vector_norm <- norm(data_constant)
print(vector_norm)
```


### Vector with NA

```{r}
vector_na <- c(1, 3, NA, 5)
```

The second param has a default option `na.rm = TRUE`, which removes NA value
```{r, error = TRUE}
vector_na_norm <- norm(vector_na)
print(vector_na_norm)
```

Apply norm() with `na.rm = FALSE` will consider NA value, then `min(vector_na)` and `max(vector_na)` are all `NA`. It's not possible to calculate norm, so we will just return original data
```{r, error = TRUE}
vector_na_norm <- norm(vector_na, FALSE)
print(vector_na_norm)
```
## List is not acceptable

List is not acceptable when apply to `min` and `max`
Below is a vector contains logical data.

```{r}
list_error <- list(1:5)
```

Apply norm() to it will produce an error:
```{r, error = TRUE}
list_error_norm <- norm(list_error)
```

